Solving the independent set problem by generic programming tensor networks \thanks{\funding{...}}

### 2021-08-25 paper/paper.tex
save

\section{Introduction}
\blue{efficient counting maximal independent sets~\cite{Manne2013}}
In this work, we introduce a tensor based framework to study the famous graph problem of finding independent sets.
Given an undirected graph $G = (V,E)$, an independent set $I \subseteq V$ is a set that for any $u,v \in I$, there is no edge connecting $u$ and $v$ in $G$.
The problem of finding the maximum independent set (MIS) size $\alpha(G) \equiv \max_{I}|I|$ belongs to the complexity class NP-complete~\cite{Hastad1996}, which is unlikely to be decided in polynomial time.
It is hard to even approximate this size in polynomial time within a factor $|V|^{1-\epsilon}$ for an arbitrarily small positive $\epsilon$.
The exhaustive search for a solution costs time $2^{|V|}$.
More efficient algorithms to compute the MIS size exactly includes the branching algorithm and dynamic programming.
Without changing the fact of exponential scaling in computing time, the branching algorithm gives a smaller base.
For example, in ~\cite{Xiao2017}, a sophisticated branching algorithm has a time complexity $1.1893^n n ^{O(1)}$.
The dynamic programming approach~\cite{Courcelle1990,Fomin2013} works better for graphs with small tree width $tw(G)$, it gives an algorithms of complexity $O(2^{tw(G)}tw(G)n)$.
People are interested in solving the independent set problem better not only because it is an NP-complete problem that directly related to other NP-complete problems like maximal cliques and vertex cover~\cite{Moore2011},
but also for its close relation with physical applications like hard spheres lattice gas model~\cite{Dyre2016}, and Rydberg hamiltonian~\cite{Pichler2018}.
However, in these applications, knowing the MIS size and one of the optimum solutions is not the only goal.
%A set is independent if and only if it is a clique in the graph's complement, so the two concepts are complementary.
%A set is independent if and only if its complement is a vertex cover. Therefore, the sum of the size of the largest independent set $\alpha (G)$ and the size of a minimum vertex cover $\beta (G)$ is equal to the number of vertices in the graph.
People often ask different questions about independent sets in order to understand the landscape of their models better.
These questions includes but not limited to, counting all independent sets, obtaining all independent sets of size $\alpha(G)$ and $\alpha(G)-1$,
counting the number of (maximal) independent sets of different sizes, and understanding the effect of a local gadget.
In this work, we attack this problem by mapping it to an generic tensor network.
It does not give a better time complexity comparing to dynamic programming, but is versatile enough to answer the above questions by engineering the tensor elements with minimum effort.

@@

>>>>

\section{Introduction}
\blue{efficient counting maximal independent sets~\cite{Manne2013}, overlap gap property~\cite{Gamarnik2013, Gamarnik2019}, independence polynomial at -1~\cite{Bousquet2008,Levit2009}}
In this work, we introduce a tensor based framework to study the famous graph problem of finding independent sets.
Given an undirected graph $G = (V,E)$, an independent set $I \subseteq V$ is a set that for any $u,v \in I$, there is no edge connecting $u$ and $v$ in $G$.
The problem of finding the maximum independent set (MIS) size $\alpha(G) \equiv \max_{I}|I|$ belongs to the complexity class NP-complete~\cite{Hastad1996}, which is unlikely to be decided in polynomial time.
It is hard to even approximate this size in polynomial time within a factor $|V|^{1-\epsilon}$ for an arbitrarily small positive $\epsilon$.
The exhaustive search for a solution costs time $2^{|V|}$.
More efficient algorithms to compute the MIS size exactly includes the branching algorithm and dynamic programming.
Without changing the fact of exponential scaling in computing time, the branching algorithm gives a smaller base.
For example, in ~\cite{Xiao2017}, a sophisticated branching algorithm has a time complexity $1.1893^n n ^{O(1)}$.
The dynamic programming approach~\cite{Courcelle1990,Fomin2013} works better for graphs with small tree width $tw(G)$, it gives an algorithms of complexity $O(2^{tw(G)}tw(G)n)$.
People are interested in solving the independent set problem better not only because it is an NP-complete problem that directly related to other NP-complete problems like maximal cliques and vertex cover~\cite{Moore2011},
but also for its close relation with physical applications like hard spheres lattice gas model~\cite{Dyre2016}, and Rydberg hamiltonian~\cite{Pichler2018}.
However, in these applications, knowing the MIS size and one of the optimum solutions is not the only goal.
%A set is independent if and only if it is a clique in the graph's complement, so the two concepts are complementary.
%A set is independent if and only if its complement is a vertex cover. Therefore, the sum of the size of the largest independent set $\alpha (G)$ and the size of a minimum vertex cover $\beta (G)$ is equal to the number of vertices in the graph.
People often ask different questions about independent sets in order to understand the landscape of their models better.
These questions includes but not limited to, counting all independent sets, obtaining all independent sets of size $\alpha(G)$ and $\alpha(G)-1$,
counting the number of (maximal) independent sets of different sizes, and understanding the effect of a local gadget.
In this work, we attack this problem by mapping it to an generic tensor network.
It does not give a better time complexity comparing to dynamic programming, but is versatile enough to answer the above questions by engineering the tensor elements with minimum effort.

@@

\section*{Acknowledgments}
We would like to acknowledge Sepehr Ebadi, Maddie Cain and Leo Zhou for popping up inspiring questions,
their questions are the driving force of this project.
Thank Chris Elord for helping us writting the fastest matrix multiplication libary for GEMM, TropicalGEMM.jl, he is a remarkable guy!
\blue{funding information}
